Internal models of type theory and internal higher categories for the Univalent Foundations

### 2014-05-23 article-models-tt.tex
tagging equations
a few more tags added
nearly finished with the section about fiber inclusions
initiated CwF section

Besides contexts, families and terms there will also be a notion of `type in
a context'. Only a family of contexts over a context $\Gamma$ is eligible to
be a type in the context $\Gamma$ and we have the intuition that `being a type'
expresses the property of `being an atomic or irreducible family of contexts'.
We will not axiomatize that every context is either judgmentally equal to the
empty context or judgmentally equal to the extension of a context by a type,
although it is certainly worth investigating the class of contexts which fall
under this category.

@@

\subsubsection{The empty context}
\label{empty}
In the theory of contexts, families and terms we introduce an empty context and,
for every context $\Gamma$, an empty family over $\Gamma$. 
When contexts are viewed as statements, the empty context is the statement that
asserts nothing. Likewise, in any context one can assert nothing and this is
achieved by the empty family. 
Since somebody who states nothing asserts it all right, there is a term of the empty
family over any context. Moreover, any two terms of the empty family are always
judgmentally equal.
\begin{align}
& \inference
  { }
  { \jctx{\emptyc}
    }
  \\
& \inference
  { \jctx{\Gamma}
    }
  { \jfam{\Gamma}{\emptyf[\Gamma]}
    }
  \\
& \inference
  { \jctx{\Gamma}
    }
  { \jterm{\Gamma}{\emptyf[\Gamma]}{\emptytm[\Gamma]}
    }
  \\
& \inference
  { \jterm{\Gamma}{\emptyf[\Gamma]}{x}
    }
  { \jtermeq
      {\Gamma}
      {\emptyf[\Gamma]}
      {x}
      {\emptytm[\Gamma]}
    }
\end{align}

@@

The weakening operation acts on three levels: on contexts, on families and
on terms. The `action on contexts' of weakening is the action we described
above: it takes a family $B$ over $\Gamma$ to a family $\ctxwk{A}{B}$ over
$\ctxext{\Gamma}{A}$; the `action on families' of weakening takes a family
$Q$ over $\ctxext{\Gamma}{B}$ to a family $\ctxwk[\famsym]{A}{Q}$ over
$\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}$; the `action on terms' of weakening takes
a term $g$ of $Q$ to a term $\ctxwk[\tmsym]{A}{g}$ of $\ctxwk[\famsym]{A}{Q}$.
\begin{align}
& \inference
  { \jfam{\Gamma}{A}
    \jfam{\Gamma}{B}
    }
  { \jfam{{\Gamma}{A}}{\ctxwk{A}{B}}
    }
& & \inference
    { \jfameq{\Gamma}{A}{A'}
      \jfameq{\Gamma}{B}{B'}
      }
    { \jfameq{{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A'}{B'}}
      }
    \\
& \inference
  { \jfam{\Gamma}{A}
    \jfam{{\Gamma}{B}}{Q}
    }
  { \jfam{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk[\famsym]{A}{Q}}
    }
& & \inference
    { \jfameq{\Gamma}{A}{A'}
      \jfameq{{\Gamma}{B}}{Q}{Q'}
      }
    { \jfameq
        {{{\Gamma}{A}}{\ctxwk{A}{B}}}
        {\ctxwk[\famsym]{A}{Q}}
        {\ctxwk[\famsym]{A'}{Q'}}
      }
    \\
%& \inference
%  { \jterm{\Gamma}{B}{y}
%    }
%  { \jterm{{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A}{y}}
%    }
%& & \inference
%    { \jtermeq{\Gamma}{B}{y}{y'}
%      }
%    { \jtermeq
%        {{\Gamma}{A}}
%        {\ctxwk{A}{B}}
%        {\ctxwk{A}{y}}
%        {\ctxwk{A'}{y'}}
%      }
%    \\
& \inference
  { \jfam{\Gamma}{A}
    \jterm{{\Gamma}{B}}{Q}{g}
    }
  { \jterm{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk[\famsym]{A}{Q}}{\ctxwk[\tmsym]{A}{g}}
    }
& & \inference
    { \jfameq{\Gamma}{A}{A'}
      \jtermeq{{\Gamma}{B}}{Q}{g}{g'}
      }
    { \jtermeq
        {{{\Gamma}{A}}{\ctxwk{A}{B}}}
        {\ctxwk[\famsym]{A}{Q}}
        {\ctxwk[\tmsym]{A}{g}}
        {\ctxwk[\tmsym]{A'}{g'}}
      }
\end{align}
The following inference rules express that when the empty context or family is
weakened, the result is the empty family.
\label{comp-w0}
\begin{align}
%& \inference
%  { \jctx{\Gamma}
%    }
%  { \jfameq{\Gamma}{\ctxwk{\Gamma}{\emptyc}}{\emptyf}
%    }
%  \\
& \inference
  { \jfam{\Gamma}{A}
    }
  { \jfameq{{\Gamma}{A}}{\ctxwk{A}{\emptyf}}{\emptyf}
    }
  \\
& \inference
  { \jfam{\Gamma}{A}
    \jfam{\Gamma}{B}
    }
  { \jfameq
    {{{\Gamma}{A}}{\ctxwk{A}{B}}}
    {\ctxwk[\famsym]{A}{\emptyf}}
    {\emptyf}
    }
\end{align}
Note that we do not have to axiomatize the action of weakening
by $A$ on the term of the empty family, because the result will be in the empty 
family and every term in the empty family is judgmentally equal to the term 
$\emptytm$.

@@

Note that we can also weaken by the empty family over $\Gamma$ itself.
Weakening by the empty family $\emptyf$ over a context $\Gamma$ leaves families, 
their terms, families over those families and terms of those unchanged:
\label{comp-0w}\begin{align}
& \inference
  { \jfam{\Gamma}{B}
    }
  { \jfameq{\Gamma}{\ctxwk{\emptyf}{B}}{B}
    }
  \\
& \inference
  { \jfam{{\Gamma}{B}}{Q}
    }
  { \jfameq{{\Gamma}{B}}{\ctxwk{\emptyf}{Q}}{Q}
    }
  \\
& \inference
  { \jterm{{\Gamma}{B}}{Q}{g}
    }
  { \jtermeq{{\Gamma}{B}}{Q}{\ctxwk{\emptyf}{g}}{g}
    }
\end{align}

\subsubsection{Substitution}
\label{substitution}
Given a family $P$ over $A$ and a term $x$ of $A$, substitution gives a way to
consider the \emph{fiber $\subst{x}{P}$ of $P$ at $x$}. As was the case with
weakening, the substitution operation comes in three parts: the `action on
contexts' of substitution is the operation just described; the `action on
families' of substitution takes a family $Q$ over $\ctxext{{\Gamma}{A}}{P}$
to a family $\subst[\famsym]{x}{Q}$ over $\ctxext{\Gamma}{\subst{x}{P}}$; the
`action on terms' of substitution takes a term $g$ of $Q$ over
$\ctxext{{\Gamma}{A}}{P}$ to a term $\subst[\tmsym]{x}{g}$ of 
$\subst[\famsym]{x}{Q}$.
\begin{align}
& \inference
  { \jterm{\Gamma}{A}{x}
    \jfam{{\Gamma}{A}}{P}
    }
  { \jfam{\Gamma}{\subst{x}{P}}
    }
& & \inference
    { \jtermeq{\Gamma}{A}{x}{x'}
      \jfameq{{\Gamma}{A}}{P}{P'}
      }
    { \jfameq{\Gamma}{\subst{x}{P}}{\subst{x'}{P'}}
      }
    \\
& \inference
  { \jterm{\Gamma}{A}{x}
    \jfam{{{\Gamma}{A}}{P}}{Q}
    }
  { \jfam{{\Gamma}{\subst{x}{P}}}{\subst[\famsym]{x}{Q}}
    }
& & \inference
    { \jtermeq{\Gamma}{A}{x}{x'}
      \jfameq{{{\Gamma}{A}}{P}}{Q}{Q'}
      }
    { \jfameq{{\Gamma}{\subst{x}{P}}}{\subst[\famsym]{x}{Q}}{\subst[\famsym]{x'}{Q'}}
      }
    \\
%& \inference
%  { \jterm{{\Gamma}{A}}{P}{f}
%    }
%  { \jterm{\Gamma}{\subst{x}{P}}{\subst{x}{f}}
%    }
%& & \inference
%    { \jtermeq{{\Gamma}{A}}{P}{f}{f'}
%      }
%    { \jtermeq{\Gamma}{\subst{x}{P}}{\subst{x}{f}}{\subst{x'}{f'}}
%      }
%    \\
& \inference
  { \jterm{\Gamma}{A}{x}
    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
    }
  { \jterm{{\Gamma}{\subst{x}{P}}}{\subst[\famsym]{x}{Q}}{\subst[\tmsym]{x}{g}}
    }
& & \inference
    { \jtermeq{\Gamma}{A}{x}{x'}
      \jtermeq{{{\Gamma}{A}}{P}}{Q}{g}{g'}
      }
    { \jtermeq
        {{\Gamma}{\subst{x}{P}}}
        {\subst[\famsym]{x}{Q}}
        {\subst[\tmsym]{x}{g}}
        {\subst[\tmsym]{x'}{g'}}
      }
\end{align}
The following inference rules establish that the fibers of the empty family are 
the empty families:
\label{comp-s0}
\begin{align}
& \inference
  { \jterm{\Gamma}{A}{x}
    }
  { \jfameq{\Gamma}{\subst{x}{\emptyf}}{\emptyf}
    }
  \\
& \inference
  { \jterm{\Gamma}{A}{x}
    \jfam{{\Gamma}{A}}{P}
    }
  { \jfameq
      {{\Gamma}{\subst{x}{P}}}
      {\subst{x}{\emptyf}}
      {\emptyf}
    }
\end{align}
Note again that we do not have to axiomatize the action of substitution by $x$ on the term
of the empty family, because the result will be in the empty family and every
term in the empty family is judgmentally equal to the term $\emptytm$.

@@

\label{comp-0s}
The following rules assert that substituting by the term $\jterm{\Gamma}{\emptyf}{\emptytm}$
leaves everything unchanged.
\begin{align}
& \inference
  { \jfam{\Gamma}{A}
    }
  { \jfameq{\Gamma}{\subst{\emptytm}{A}}{A}
    }
  \\
%& \inference
%  { \jterm{\Gamma}{A}{x}
%    }
%  { \jtermeq{\Gamma}{A}{\subst{\emptytm}{x}}{x}
%    }
%  \\
& \inference
  { \jfam{{\Gamma}{A}}{P}
    }
  { \jfameq{{\Gamma}{A}}{\subst{\emptytm}{P}}{P}
    }
  \\
& \inference
  { \jterm{{\Gamma}{A}}{P}{f}
    }
  { \jtermeq{{\Gamma}{A}}{P}{\subst{\emptytm}{f}}{f}
    }.
\end{align}

@@

\subsubsection{The identity term of a weakened family}
\label{comp-wi}
The identity term of a weakened family is the weakened identity term:
\begin{equation}\label{idfunc-wk-comp}
\inference
  { \jfam{\Gamma}{A}
    \jfam{\Gamma}{B}
    }
  { \unfoldall{\jhomeq
      {{\Gamma}{A}}
      {\ctxwk{A}{B}}
      {\ctxwk{A}{B}}
      {\ctxwk{A}{\idtm{B}}}
      {\idtm{\ctxwk{A}{B}}}
    }}
\end{equation}

\subsubsection{The identity term of a substituted family}
\label{comp-si}
The identity term of a substituted family is the substitution of the identity term
\begin{equation}\label{idfunc-subst-comp}
\inference
  { \jterm{\Gamma}{A}{x}
    \jfam{{\Gamma}{A}}{P}
    }
  { \unfoldall{\jhomeq
      {\Gamma}
      {\subst{x}{P}}
      {\subst{x}{P}}
      {\subst{x}{\idtm{P}}}
      {\idtm{\subst{x}{P}}}
    }}
\end{equation}

@@

\subsubsection{The cancellation property of weakening and substitution}
\label{cancellation-ws}
The judgmental equalities we're about to describe assert that substituting a term
in the weakening a thing gives you the thing back. In the case of contexts we get that each fiber
$\subst{x}{\ctxwk{A}{B}}$ is just $B$ and in the case of terms we get 
that $\ctxwk{A}{y}$ is the constant function
mapping everything to $y:B$. Thus, these rules actually establish the weakening
as the weakening. After stating the rules we will describe what it means to
compose context morphisms (terms of weakened contexts).
\begin{align}
& \inference
  { \jfam{\Gamma}{A}
    \jfam{{\Gamma}{B}}{Q}
    \jterm{\Gamma}{A}{x}
    }
  { \jfameq{{\Gamma}{B}}{\subst{x}{\ctxwk{A}{Q}}}{Q}
    }
  \label{defn-ws-3}\\
& \inference
  { \jterm{\Gamma}{A}{x}
    \jterm{{\Gamma}{B}}{Q}{g}
    }
  { \jtermeq{{\Gamma}{B}}{Q}{\subst{x}{\ctxwk{A}{g}}}{g}
    }
  \label{defn-ws-4}
\end{align}

\subsubsection{The cancellation properties of identity terms}
\label{cancellation-i}
Identity terms are determined by their behavior with respect to substitution combined with
weakening. The identity terms will also be subject to compatibility rules.
\begin{align}
& \inference
  { \jterm{\Gamma}{A}{x}
    }
  { \jtermeq{\Gamma}{A}{\subst{x}{\idtm{A}}}{x}
    }
  \label{idfunc-subst-defn}\\
& \inference
  { \jfam{{\Gamma}{A}}{P}
    }
  { \jfameq{{\Gamma}{A}}{\subst{\idtm{A}}{\ctxwk{A}{P}}}{P}
    }
  \label{idfunc-wk-defn}\\
%An unnecessary rule:
%& \inference
%  { \jfam{{\Gamma}{A}}{P}
%    }
%  { \jfameq
%      {{\Gamma}{A}}
%      {\subst{\idtm{A}}{\ctxwk{{A}{A}}{P}}}
%      {P}
%    }
%  \label{idfunc-wk-defn2}\\
& \inference
  { \jterm{{\Gamma}{A}}{P}{f}
    }
  { \jtermeq
      {{\Gamma}{A}}
      {P}
      {\subst{\idtm{A}}{\ctxwk{A}{f}}}
      {f}
    }
  \label{idfunc-precomp}\\
%An unnecessary rule:
%& \inference
%  { \jterm{{\Gamma}{A}}{P}{f}
%    }
%  { \jtermeq
%      {{\Gamma}{A}}
%      {P}
%      {\subst{\idtm{A}}{\ctxwk{{A}{A}}{f}}}
%      {f}
%    }
%  \label{idfunc-precomp}\\
& \inference
  { \unfoldall{\jhom{\Gamma}{A}{B}{f}}
    }
  { \unfoldall{\jhomeq{\Gamma}{A}{B}{\jcomp{A}{f}{\idtm{B}}}{f}}
    }
  \label{idfunc-postcomp}
\end{align}

>>>>

Besides contexts, families and terms there will also be a notion of `type in
a context', which one may assume.
Only a family of contexts over a context $\Gamma$ is eligible to
be a type in the context $\Gamma$ and we have the intuition that `being a type'
expresses the property of `being an atomic or irreducible family of contexts'.
We will not axiomatize that every context is either judgmentally equal to the
empty context or judgmentally equal to the extension of a context by a type,
although it is certainly worth investigating the class of contexts which fall
under this category. However, we will primarily investigate families of
contexts and leave discussion on the possibility of a typing judgment until the
end of this section. With families of contexts being the principal things of
study, we have added a bit more generality to the theory, compared to the
theory of categories with families presented in \cite{Dybjer1996}. The reason
for this added generality is twofold:
\begin{enumerate}
\item We want to be very sure that we can interpret all the rules of type theory
in a straightforward manner in any context.
\item We do want do end up with a categorical structure on the contexts, derived
from the theory by requiring that a morphism from $\Gamma$ to $\Delta$ is a
term of the weakening of $\Delta$ by $\Gamma$ in context $\Gamma$. Thus, we
must be able to lift the context $\Delta$ to become a family of contexts over
$\Gamma$ and it is not reasonable to expect that this results in a type.
\end{enumerate}

@@

\subsubsection{The empty context}
\label{empty}
In the theory of contexts, families and terms we introduce an empty context and,
for every context $\Gamma$, an empty family over $\Gamma$. 
When contexts are viewed as statements, the empty context is the statement that
asserts nothing. Likewise, in any context one can assert nothing and this is
achieved by the empty family. 
Since somebody who states nothing asserts it all right, there is a term of the empty
family over any context. Moreover, any two terms of the empty family are always
judgmentally equal.
\begin{align}
& \inference
  { }
  { \jctx{\emptyc}
    }
  \\
& \inference
  { \jctx{\Gamma}
    }
  { \jfam{\Gamma}{\emptyf[\Gamma]}
    }
  \\
& \inference
  { \jctx{\Gamma}
    }
  { \jterm{\Gamma}{\emptyf[\Gamma]}{\emptytm[\Gamma]}
    }
  \\
& \inference
  { \jterm{\Gamma}{\emptyf[\Gamma]}{x}
    }
  { \jtermeq
      {\Gamma}
      {\emptyf[\Gamma]}
      {x}
      {\emptytm[\Gamma]}
    }
  \label{comp-00-t}
\end{align}

@@

The weakening operation acts on three levels: on contexts, on families and
on terms. The `action on contexts' of weakening is the action we described
above: it takes a family $B$ over $\Gamma$ to a family $\ctxwk{A}{B}$ over
$\ctxext{\Gamma}{A}$; the `action on families' of weakening takes a family
$Q$ over $\ctxext{\Gamma}{B}$ to a family $\ctxwk[\famsym]{A}{Q}$ over
$\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}$; the `action on terms' of weakening takes
a term $g$ of $Q$ to a term $\ctxwk[\tmsym]{A}{g}$ of $\ctxwk[\famsym]{A}{Q}$.
\begin{align}
& \inference
  { \jfam{\Gamma}{A}
    \jfam{\Gamma}{B}
    }
  { \jfam{{\Gamma}{A}}{\ctxwk{A}{B}}
    }
& & \inference
    { \jfameq{\Gamma}{A}{A'}
      \jfameq{\Gamma}{B}{B'}
      }
    { \jfameq{{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A'}{B'}}
      }
    \\
& \inference
  { \jfam{\Gamma}{A}
    \jfam{{\Gamma}{B}}{Q}
    }
  { \jfam{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk[\famsym]{A}{Q}}
    }
& & \inference
    { \jfameq{\Gamma}{A}{A'}
      \jfameq{{\Gamma}{B}}{Q}{Q'}
      }
    { \jfameq
        {{{\Gamma}{A}}{\ctxwk{A}{B}}}
        {\ctxwk[\famsym]{A}{Q}}
        {\ctxwk[\famsym]{A'}{Q'}}
      }
    \\
%& \inference
%  { \jterm{\Gamma}{B}{y}
%    }
%  { \jterm{{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A}{y}}
%    }
%& & \inference
%    { \jtermeq{\Gamma}{B}{y}{y'}
%      }
%    { \jtermeq
%        {{\Gamma}{A}}
%        {\ctxwk{A}{B}}
%        {\ctxwk{A}{y}}
%        {\ctxwk{A'}{y'}}
%      }
%    \\
& \inference
  { \jfam{\Gamma}{A}
    \jterm{{\Gamma}{B}}{Q}{g}
    }
  { \jterm{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk[\famsym]{A}{Q}}{\ctxwk[\tmsym]{A}{g}}
    }
& & \inference
    { \jfameq{\Gamma}{A}{A'}
      \jtermeq{{\Gamma}{B}}{Q}{g}{g'}
      }
    { \jtermeq
        {{{\Gamma}{A}}{\ctxwk{A}{B}}}
        {\ctxwk[\famsym]{A}{Q}}
        {\ctxwk[\tmsym]{A}{g}}
        {\ctxwk[\tmsym]{A'}{g'}}
      }
\end{align}
The following inference rules express that when the empty context or family is
weakened, the result is the empty family.
\label{comp-w0}
\begin{align}
%& \inference
%  { \jctx{\Gamma}
%    }
%  { \jfameq{\Gamma}{\ctxwk{\Gamma}{\emptyc}}{\emptyf}
%    }
%  \\
& \inference
  { \jfam{\Gamma}{A}
    }
  { \jfameq{{\Gamma}{A}}{\ctxwk{A}{\emptyf}}{\emptyf}
    }
  \label{comp-w0-c}\\
& \inference
  { \jfam{\Gamma}{A}
    \jfam{\Gamma}{B}
    }
  { \jfameq
    {{{\Gamma}{A}}{\ctxwk{A}{B}}}
    {\ctxwk[\famsym]{A}{\emptyf}}
    {\emptyf}
    }
  \label{comp-w0-f}
\end{align}
Note that we do not have to axiomatize the action of weakening
by $A$ on the term of the empty family, because the result will be in the empty 
family and every term in the empty family is judgmentally equal to the term 
$\emptytm$.

@@

Note that we can also weaken by the empty family over $\Gamma$ itself.
Weakening by the empty family $\emptyf$ over a context $\Gamma$ leaves families, 
their terms, families over those families and terms of those unchanged:
\label{comp-0w}\begin{align}
& \inference
  { \jfam{\Gamma}{B}
    }
  { \jfameq{\Gamma}{\ctxwk{\emptyf}{B}}{B}
    }
  \label{comp-0w-c}
  \\
& \inference
  { \jfam{{\Gamma}{B}}{Q}
    }
  { \jfameq{{\Gamma}{B}}{\ctxwk{\emptyf}{Q}}{Q}
    }
  \label{comp-0w-f}
  \\
& \inference
  { \jterm{{\Gamma}{B}}{Q}{g}
    }
  { \jtermeq{{\Gamma}{B}}{Q}{\ctxwk{\emptyf}{g}}{g}
    }
  \label{comp-0w-t}
\end{align}

\subsubsection{Substitution}
\label{substitution}
Given a family $P$ over $A$ and a term $x$ of $A$, substitution gives a way to
consider the \emph{fiber $\subst{x}{P}$ of $P$ at $x$}. As was the case with
weakening, the substitution operation comes in three parts: the `action on
contexts' of substitution is the operation just described; the `action on
families' of substitution takes a family $Q$ over $\ctxext{{\Gamma}{A}}{P}$
to a family $\subst[\famsym]{x}{Q}$ over $\ctxext{\Gamma}{\subst{x}{P}}$; the
`action on terms' of substitution takes a term $g$ of $Q$ over
$\ctxext{{\Gamma}{A}}{P}$ to a term $\subst[\tmsym]{x}{g}$ of 
$\subst[\famsym]{x}{Q}$.
\begin{align}
& \inference
  { \jterm{\Gamma}{A}{x}
    \jfam{{\Gamma}{A}}{P}
    }
  { \jfam{\Gamma}{\subst{x}{P}}
    }
& & \inference
    { \jtermeq{\Gamma}{A}{x}{x'}
      \jfameq{{\Gamma}{A}}{P}{P'}
      }
    { \jfameq{\Gamma}{\subst{x}{P}}{\subst{x'}{P'}}
      }
    \\
& \inference
  { \jterm{\Gamma}{A}{x}
    \jfam{{{\Gamma}{A}}{P}}{Q}
    }
  { \jfam{{\Gamma}{\subst{x}{P}}}{\subst[\famsym]{x}{Q}}
    }
& & \inference
    { \jtermeq{\Gamma}{A}{x}{x'}
      \jfameq{{{\Gamma}{A}}{P}}{Q}{Q'}
      }
    { \jfameq{{\Gamma}{\subst{x}{P}}}{\subst[\famsym]{x}{Q}}{\subst[\famsym]{x'}{Q'}}
      }
    \\
%& \inference
%  { \jterm{{\Gamma}{A}}{P}{f}
%    }
%  { \jterm{\Gamma}{\subst{x}{P}}{\subst{x}{f}}
%    }
%& & \inference
%    { \jtermeq{{\Gamma}{A}}{P}{f}{f'}
%      }
%    { \jtermeq{\Gamma}{\subst{x}{P}}{\subst{x}{f}}{\subst{x'}{f'}}
%      }
%    \\
& \inference
  { \jterm{\Gamma}{A}{x}
    \jterm{{{\Gamma}{A}}{P}}{Q}{g}
    }
  { \jterm{{\Gamma}{\subst{x}{P}}}{\subst[\famsym]{x}{Q}}{\subst[\tmsym]{x}{g}}
    }
& & \inference
    { \jtermeq{\Gamma}{A}{x}{x'}
      \jtermeq{{{\Gamma}{A}}{P}}{Q}{g}{g'}
      }
    { \jtermeq
        {{\Gamma}{\subst{x}{P}}}
        {\subst[\famsym]{x}{Q}}
        {\subst[\tmsym]{x}{g}}
        {\subst[\tmsym]{x'}{g'}}
      }
\end{align}
The following inference rules establish that the fibers of the empty family are 
the empty families:
\label{comp-s0}
\begin{align}
& \inference
  { \jterm{\Gamma}{A}{x}
    }
  { \jfameq{\Gamma}{\subst{x}{\emptyf}}{\emptyf}
    }
  \label{comp-s0-c}
  \\
& \inference
  { \jterm{\Gamma}{A}{x}
    \jfam{{\Gamma}{A}}{P}
    }
  { \jfameq
      {{\Gamma}{\subst{x}{P}}}
      {\subst{x}{\emptyf}}
      {\emptyf}
    }
  \label{comp-s0-f}
\end{align}
Note again that we do not have to axiomatize the action of substitution by $x$ on the term
of the empty family, because the result will be in the empty family and every
term in the empty family is judgmentally equal to the term $\emptytm$.

@@

\label{comp-0s}
The following rules assert that substituting by the term $\jterm{\Gamma}{\emptyf}{\emptytm}$
leaves everything unchanged.
\begin{align}
& \inference
  { \jfam{\Gamma}{A}
    }
  { \jfameq{\Gamma}{\subst{\emptytm}{A}}{A}
    }
  \label{comp-0s-c}
  \\
%& \inference
%  { \jterm{\Gamma}{A}{x}
%    }
%  { \jtermeq{\Gamma}{A}{\subst{\emptytm}{x}}{x}
%    }
%  \\
& \inference
  { \jfam{{\Gamma}{A}}{P}
    }
  { \jfameq{{\Gamma}{A}}{\subst{\emptytm}{P}}{P}
    }
  \label{comp-0s-f}
  \\
& \inference
  { \jterm{{\Gamma}{A}}{P}{f}
    }
  { \jtermeq{{\Gamma}{A}}{P}{\subst{\emptytm}{f}}{f}
    }.
  \label{comp-0s-t}
\end{align}

@@

\subsubsection{The identity term of a weakened family}
\label{comp-wi}
The identity term of a weakened family is the weakened identity term:
\begin{equation}
\inference
  { \jfam{\Gamma}{A}
    \jfam{\Gamma}{B}
    }
  { \unfoldall{\jhomeq
      {{\Gamma}{A}}
      {\ctxwk{A}{B}}
      {\ctxwk{A}{B}}
      {\ctxwk{A}{\idtm{B}}}
      {\idtm{\ctxwk{A}{B}}}
    }}
  \label{comp-wi-t}
\end{equation}

\subsubsection{The identity term of a substituted family}
\label{comp-si}
The identity term of a substituted family is the substitution of the identity term
\begin{equation}
\inference
  { \jterm{\Gamma}{A}{x}
    \jfam{{\Gamma}{A}}{P}
    }
  { \unfoldall{\jhomeq
      {\Gamma}
      {\subst{x}{P}}
      {\subst{x}{P}}
      {\subst{x}{\idtm{P}}}
      {\idtm{\subst{x}{P}}}
    }}
  \label{comp-si-t}
\end{equation}

@@

\subsubsection{The cancellation property of weakening and substitution}
\label{cancellation-ws}
The judgmental equalities we're about to describe assert that substituting a term
in the weakening a thing gives you the thing back. In the case of contexts we get that each fiber
$\subst{x}{\ctxwk{A}{B}}$ is just $B$ and in the case of terms we get 
that $\ctxwk{A}{y}$ is the constant function
mapping everything to $y:B$. Thus, these rules actually establish the weakening
as the weakening. After stating the rules we will describe what it means to
compose context morphisms (terms of weakened contexts).
\begin{align}
& \inference
  { \jfam{\Gamma}{A}
    \jfam{{\Gamma}{B}}{Q}
    \jterm{\Gamma}{A}{x}
    }
  { \jfameq{{\Gamma}{B}}{\subst{x}{\ctxwk{A}{Q}}}{Q}
    }
  \label{cancellation-ws-f}\\
& \inference
  { \jterm{\Gamma}{A}{x}
    \jterm{{\Gamma}{B}}{Q}{g}
    }
  { \jtermeq{{\Gamma}{B}}{Q}{\subst{x}{\ctxwk{A}{g}}}{g}
    }
  \label{cancellation-ws-t}
\end{align}

\subsubsection{The cancellation properties of identity terms}
\label{cancellation-i}
Identity terms are determined by their behavior with respect to substitution combined with
weakening. The identity terms will also be subject to compatibility rules.
\begin{align}
& \inference
  { \jterm{\Gamma}{A}{x}
    }
  { \jtermeq{\Gamma}{A}{\subst{x}{\idtm{A}}}{x}
    }
  \label{cancellation-si}\\
& \inference
  { \jfam{{\Gamma}{A}}{P}
    }
  { \jfameq{{\Gamma}{A}}{\subst{\idtm{A}}{\ctxwk{A}{P}}}{P}
    }
  \label{idfunc-wk-defn}\\
%An unnecessary rule:
%& \inference
%  { \jfam{{\Gamma}{A}}{P}
%    }
%  { \jfameq
%      {{\Gamma}{A}}
%      {\subst{\idtm{A}}{\ctxwk{{A}{A}}{P}}}
%      {P}
%    }
%  \label{idfunc-wk-defn2}\\
& \inference
  { \jterm{{\Gamma}{A}}{P}{f}
    }
  { \jtermeq
      {{\Gamma}{A}}
      {P}
      {\subst{\idtm{A}}{\ctxwk{A}{f}}}
      {f}
    }
  \label{idfunc-precomp}\\
%An unnecessary rule:
%& \inference
%  { \jterm{{\Gamma}{A}}{P}{f}
%    }
%  { \jtermeq
%      {{\Gamma}{A}}
%      {P}
%      {\subst{\idtm{A}}{\ctxwk{{A}{A}}{f}}}
%      {f}
%    }
%  \label{idfunc-precomp}\\
& \inference
  { \unfoldall{\jhom{\Gamma}{A}{B}{f}}
    }
  { \unfoldall{\jhomeq{\Gamma}{A}{B}{\jcomp{A}{f}{\idtm{B}}}{f}}
    }
  \label{idfunc-postcomp}
\end{align}
